---
title: "Bone Marrow Transplant Data Set"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{bone_Marrow_Transplant_Data}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```


## Introduction

We apply the package __ensBagg__ to a data set The Bone Marrow Transplant Data (bmt) which is in the package __timereg__ . Bone marrow transplants are a standard treatment for acute leukemia. In the recovery process patients can suffer from Infections, toxicity, and (after allogeneic HSCT only), graft-vs.-host disease (GVHD) being  the main causes of death. We are interested in predicting  the time to death related to treatment where the risk of relapse is a competing risk. 

## Setup

We load the package __ensBagg__ and __timereg__ and we load the data set bmt

```{r}
library(ensBagg)
library(timereg)
bmt.data <- data(bmt)
```


````{r}
summary(bmt)
dim(bmt)
train.set <- sample(1:nrow(bmt), floor(.9*nrow(bmt)), replace=FALSE) 
test.set <- setdiff(1:nrow(bmt), train.set)
bmt.train <- data.frame(bmt[train.set,])
bmt.test <- data.frame(bmt[test.set,]) 
tao=7
````

Let's define the library of algorithms:
````{r, warning=FALSE}
ens.library <-ensBagg::ens.all.algorithms()
````

We set the covariates that we include in the models:

````{r, warning=FALSE}
xnam <- names(bmt)[-(1:2)]
xnam
````

Now, we are ready to predict the outcome of interest using all the algorithms:

````{r, warning=FALSE}
pred.bmt=ensBagg::ensBagg(train.data = bmt.train,test.data = bmt.test,xnam=xnam,tao=7,weighting ="CoxPH" ,folds =5,ens.library = ens.library )
````

The assessment of predictive performance using the IPCW AUC is:

````{r}
pred.bmt$auc_ipcwBagg
````

Now let s take a look at prediction of the algorithms that allows for weights natively:

````{r}
head(pred.bmt$prediction_native_weights,5)
````

and their performance is:

````{r}
pred.bmt$auc_native_weights
````

The prediction of the survival based methods
````{r}
head(pred.bmt$prediction_survival,5)

````
````{r}
pred.bmt$auc_survival
````

Lastly, we could see the performance of the algorithms if we were to discard the censored observations

````{r,warning=FALSE}
pred.discard <- ensBagg::prediction_discard(train.data = bmt.train,test.data = bmt.test,xnam=names(bmt)[-(1:2)],tao=7,ens.library=ens.library)
head(pred.discard$prediction_discard)
pred.discard$auc_discard
````


The ensemble ROC curve is
````{r}
ensBagg::plot_roc(time=bmt.test$time,delta = bmt.test$cause,marker =pred.bmt$prediction_ensBagg[,"Ensemble"],wts=pred.bmt$wts_test,tao=7,method = "ipcw")
````


The Random Forest ROC curve is
````{r}
ensBagg::plot_roc(time=bmt.test$time,delta = bmt.test$cause,marker =pred.bmt$prediction_ensBagg[,"ens.randomForest"],wts=pred.bmt$wts_test,tao=7,method = "ipcw")
````


The Random Forest survival ROC curve is
````{r}
ensBagg::plot_roc(time=bmt.test$time,delta = bmt.test$cause,marker =pred.bmt$prediction_survival[,"Random Forest"],wts=pred.bmt$wts_test,tao=7,method = "ipcw")
````


The Random Forest natively weighted ROC curve is
````{r}
ensBagg::plot_roc(time=bmt.test$time,delta = bmt.test$cause,marker =pred.bmt$prediction_native_weights[,"ens.randomForest"],wts=pred.bmt$wts_test,tao=7,method = "ipcw")
````


The Random Forest discarding censored observations
````{r}
ensBagg::plot_roc(time=bmt.test$time,delta = bmt.test$cause,marker =pred.discard$prediction_discard[,"ens.randomForest"],tao=7,method = "discard")
````

