---
title: "Introduction to ensBagg package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ensBagg package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



## Introduction

This vignette focuses on 


## Setup

We load the package __ensBagg__ 

```{r, eval=FALSE}
library(ensBagg)
```

## Quick start

We give a quick overview of the main functions. We will use the simulated data which ships with the package and can be loaded via:

```{r}
library(ensBagg)
data("exampleData")
```

__exampleData__ is a list where the first elemenent contains the training data set and the second element contains the test data set. 

```{r}
train <- exampleData[[1]][[1]]
head(train)
test <- exampleData[[2]][[1]]
head(test)
```

There is a third element __exampleData__ which is the true AUC for the simulated data  on the cumulative incidence of the
main event at time 26.5 computed analytically under the Weibull distribution using the scale and shape parameter for both event times.

```{r}
auc_true <- exampleData[[3]][[1]]
auc_true
```




