---
title: "Introduction to ensBagg package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction to ensBagg package}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```



## Introduction

This vignette focuses on giving an overview of the main functions of the package. This vignette uses the simulated data which ships with the package.  


## Setup

We load the package __ensBagg__ 

```{r, eval=FALSE}
library(ensBagg)
```

## Example: Simulated Data

The simulated data can be loaded via:

```{r}
library(ensBagg)
data("exampleData")
```

__exampleData__ is a list with three elements. The first elemenent contains the training data set and the second element contains the test data set: 

```{r}
train <- exampleData[[1]][[1]]
head(train)
test <- exampleData[[2]][[1]]
head(test)
```

There is a third element __exampleData__ which is the true AUC for the simulated data  on the cumulative incidence of the
main event at time 26.5 computed analytically under the Weibull distribution using the scale and shape parameter for both event times.

```{r}
auc_true <- exampleData[[3]][[1]]
auc_true
```

The train data set consists in 800 individuals where 167 of them are censored:
```{r}
summary(train)
```

The test data set consists in 200 individuals where 45 of them are censored:
```{r}
summary(train)
```


